---
title: "Title of Report"
date: last-modified
date-format: "MMMM D, YYYY"
author: 
    - name: Warren Woo
      affil-id: 1
    - name: Zachary Lakin
      affil-id: 2
    - name: Daniel Tsemekhman
      affil-id: 1
    - name: Fourth Author
      affil-id: 1
affiliations:
    - id: 1
      name: Department of Statistics, Cal Poly - SLO
abstract: ""
format: 
  pdf:
    number-sections: true
    template-partials: 
      - title.tex
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
execute: 
  warning: false
  message: false
  echo: false
bibliography: references.bib
---

```{r}
#| label: setup
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r}
#| label: data-cleaning
# this is where you might put some data cleaning steps
```

```{r}
#| label: data-analysis
# you might choose to run your analysis here or in a separate script. Save your output for later.
```

## Introduction {#intro}

Have you ever wondered what factors make certain X-Men members appear more frequently than others in the comic book series? This report aims to uncover the key variables that influence the total appearance count of X-Men characters. By analyzing a dataset of 26 X-Men members featured in issues from the 1960s to the 1990s, we can gain insights into what drives a character's prominence and staying power in this iconic franchise.

The dataset used in this analysis was compiled by Anderson Evans, a writer for the online website Rally. Evans collected the data from the Marvel Fandom Database, meticulously counting the appearances of each X-Men character in issues released between 1963 and 1992. The dataset includes 44 variables for each of the 26 characters, such as their total appearances, appearance percentages by decade, estimated monetary value, and gender.

To determine the factors that significantly impact an X-Men character's total appearances, several statistical methods will be employed. A chi-square goodness of fit test will assess if the distribution of issues per member follows a specific pattern. Linear regression will be used to model the relationship between the number of issues and other quantitative variables, such as the decade of publication. Finally, analysis of variance (ANOVA) will compare the mean number of issues across different categories, like the X-Men eras.

It is important to acknowledge some limitations of the dataset that may affect the conclusions drawn from this analysis. The unequal representation of male and female characters necessitates using appearance rates instead of raw appearance counts when examining gender differences. Additionally, the dataset includes only a subset of all X-Men characters and is limited to comics published from the 1960s to the 1990s, potentially restricting the generalizability of the findings.

This report will begin by providing an overview of the dataset and variables, followed by a detailed description of the analytical methods used. The results section will present the findings from each statistical test, and the discussion will interpret these results in the context of the research question. Finally, the conclusion will summarize the key takeaways and suggest future directions for research on factors influencing X-Men character appearances.


## Data and Methods

### Data Collection

### Statistical Analysis

## Results {#results}

You could use in line code to provide results such as: Information about `r nrow(penguins)` penguins were collected and made available by Dr. Kristen Gorman and the Palmer Station, Antarctica LTER, a member of the Long Term Ecological Research Network.

As shown in @fig-penguin-plot, we can see abc. See <https://quarto.org/docs/authoring/cross-references.html> for more about cross referencing plots.

Note! To cross reference a figure, the label must start with fig-xxx.

```{r}
#| label: fig-penguin-plot
#| fig-cap: "Your figure caption."
#| fig-alt: "Your figure alternative text."
#| fig-align: center
#| fig-pos: "H"


```
Supported by @tbl-penguin-summary, we can see abc. See <https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf> for more about styling tables.

Note! To cross reference a table, the label must start with tbl-xxx.

```{r}
#| label: tbl-penguin-summary
#| tbl-cap: "Put caption here."
#| tbl-pos: "H"

palmerpenguins::penguins |> 
  group_by(species) |> 
  summarize(avg_bill_length = mean(bill_length_mm, na.rm = T),
            sd_bill_length = sd(bill_length_mm, na.rm = T)
            ) |> 
  kableExtra::kbl(digits = 1,
                  col.names = c("Species", 
                                "Average Bill Length (mm)",
                                "Standard Deviation of Bill Length (mm)"),
                  booktabs = TRUE
                  )
```

## Discussion {#discussion}
Our report aimed to uncover the key variables that influence the total appearance count of X-Men characters. After conducting our analysis, we learned that the more profitable an X-Men character is, the more they're going to appear in X-Men comic-book issues. 
Did the data behave as you expected?
The data did not behave as we expected. 
Is the result statistically significant? How large is the effect? Would you consider it practically significant? How are you deciding? (Do you need to bring in any subject-matter knowledge here? For example 2 goals in hockey is a lot, but 2 points in basketball is a close game!)
Strengths (don’t restate but interpret) of data collection and analysis methods
Can you draw cause-and-effect conclusions?
To what population are you willing to generalize?
We can generalize our results to all X-Men characters that appeared in comic book issues in the decades between 60's and 90's.
How does your study contribute to the existing knowledge base and the larger context?
Critically assess any limitations or potential biases and how they impacted the interpretation of your findings.
Note any deficiencies or unexplained data.
Any issues with the analysis methods used?
Any other potential sources of error or biases?
 
## Conclusion {#conclusion}

## References {.unnumbered}
