---
title: "Title of Report"
date: last-modified
date-format: "MMMM D, YYYY"
author: 
    - name: Warren Woo
      affil-id: 1
    - name: Zachary Lakin
      affil-id: 2
    - name: Daniel Tsemekhman
      affil-id: 1
    - name: Fourth Author
      affil-id: 1
affiliations:
    - id: 1
      name: Department of Statistics, Cal Poly - SLO
abstract: ""
format: 
  pdf:
    number-sections: true
    template-partials: 
      - title.tex
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
execute: 
  warning: false
  message: false
  echo: false
bibliography: references.bib
---

```{r}
#| label: setup
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r}
#| label: data-cleaning
# this is where you might put some data cleaning steps
```

```{r}
#| label: data-analysis
# you might choose to run your analysis here or in a separate script. Save your output for later.
```

## Introduction {#intro}

Have you ever wondered what factors make certain X-Men members appear more frequently than others in the comic book series? This report aims to uncover the key variables that influence the total appearance count of X-Men characters. By analyzing a dataset of 26 X-Men members featured in issues from the 1960s to the 1990s, we can gain insights into what drives a character's prominence and staying power in this iconic franchise.

The dataset used in this analysis was compiled by Anderson Evans, a writer for the online website Rally. Evans collected the data from the Marvel Fandom Database, meticulously counting the appearances of each X-Men character in issues released between 1963 and 1992. The dataset includes 44 variables for each of the 26 characters, such as their total appearances, appearance percentages by decade, estimated monetary value, and gender.

To determine the factors that significantly impact an X-Men character's total appearances, several statistical methods will be employed. A chi-square goodness of fit test will assess if the distribution of issues per member follows a specific pattern. Linear regression will be used to model the relationship between the number of issues and other quantitative variables, such as the decade of publication. Finally, analysis of variance (ANOVA) will compare the mean number of issues across different categories, like the X-Men eras.

It is important to acknowledge some limitations of the dataset that may affect the conclusions drawn from this analysis. The unequal representation of male and female characters necessitates using appearance rates instead of raw appearance counts when examining gender differences. Additionally, the dataset includes only a subset of all X-Men characters and is limited to comics published from the 1960s to the 1990s, potentially restricting the generalizability of the findings.

This report will begin by providing an overview of the dataset and variables, followed by a detailed description of the analytical methods used. The results section will present the findings from each statistical test, and the discussion will interpret these results in the context of the research question. Finally, the conclusion will summarize the key takeaways and suggest future directions for research on factors influencing X-Men character appearances.



## Data and Methods


### Data Collection
The data collection took place in 2022 by Anderson Evans for his article on the website Rallyrd.com. Rally is an asset investment company which allows users to buy and sell equity shares in collectible assets. The data collection took place at the Rally company, looking at websites such as UncannyXMen.net and the Marvel Fandom Database. There are only 26 datapoints/observational units in this dataset, and it is 26 indidivual X-men. A computer was computer was used to collect the data possibly from api, and webscraping. Variables used, 
TotalValue_heritage	Total value of each X-Men team member's total number of issues as reflected by Heritage highest sale. Along with their values for 60s, 70s, 80s, and 90s. I also used the ebay counterpart for this which was the Total value of each X-Men team member's total number of issues as reflected by Ebays highest sale, and I combined these into TotalValue, and TotalValue60s, TotalValue70s, TotalValue80s, and TotalValue90s. Other variables used were the Members themselves, and Total Issues which is the total amount of issues from the 60s-90s. We are interested in if what factors influence a characters appearance rate.


### Statistical Analysis
```{r}
MMO|>
  mutate(TotalValue_heritage = TotalValue_heritage+TotalValue_ebay)


IssuesVVal <- lm(TotalIssues~ TotalValue60s_heritage + TotalValue70s_heritage  + TotalValue80s_heritage  + TotalValue90s_heritage , data = MMO)

summary(IssuesVVal)

hist(resid(IssuesVVal))

plot(IssuesVVal)



```
After running a linear regression and checking model assumptions it seems like all the FINE assumptions are met normality is good, equal varince seems okay, and independence is met through the assumption that the the values of the comics dont effect each other, but there is a light issue as the sample size is small, it is only 26 units so.


```{r}

full_model <- lm( TotalIssues~ TotalValue60s_heritage + TotalValue70s_heritage  + TotalValue80s_heritage  + TotalValue90s_heritage, data = MMO)


reduced_model <- lm( TotalIssues~ TotalValue60s_heritage + TotalValue80s_heritage+TotalValue90s_heritage, data = MMO)


partial_f_test <- anova(reduced_model, full_model)


print(partial_f_test)

```
The partial F-test assesses whether the additional variable, TotalValue70s_heritage, significantly improves the fit of the model.

The results indicate that Model 2, which includes TotalValue70s_heritage, provides a significantly better fit to the data compared to Model 1, as evidenced by the F-statistic of 4.2273 with a corresponding p-value of 0.05242. However, this p-value is marginally above the conventional significance level of 0.05, suggesting that the improvement in fit may not be statistically significant at the 5% level but could be considered significant at a higher significance level.

Therefore, while there is some evidence to suggest that TotalValue70s_heritage contributes to explaining the variability in TotalIssues, further investigation or a larger sample size may be warranted to confirm its significance. But for now I will drop the variable.

```{r}






## Results {#results}

df_stacked <- MMO %>%
  gather(key = "Period", value = "TotalValstack",TotalValue80s_heritage , TotalValue90s_heritage,TotalValue60s_heritage  )


sum1= lm(TotalValue_heritage~TotalValstack, data = df_stacked)
summary(sum1)


df_stacked|>
  ggplot(aes(y=TotalValstack, x=TotalIssues, color=Period))+
  geom_point()+
  geom_smooth(method = 'lm', se = FALSE)

```

Red data points (60s heritage) exhibit a trend with a noticeable positive slope, indicating that "TotalValsstack" increases significantly with an increase in "TotalIssues."
Green data points (80s heritage) show a slight positive trend, but the increase in "TotalValsstack" with "TotalIssues" is much less pronounced compared to the 60s heritage.
Blue data points (90s heritage) have a trend line almost flat, suggesting little to no correlation between "TotalValsstack" and "TotalIssues."

In the end, it does seem like the price of the X-men comics does predict the appearance of certain heritage periods. The linear regression analysis indicates a significant positive relationship between the total value of X-men comics and the heritage periods, as evidenced by the coefficient estimate for TotalValstack (0.8722, p < 0.001). This suggests that as the total value of X-men comics increases, there is a corresponding increase in the likelihood of certain heritage periods being represented. However, it's important to note that while this relationship is statistically significant, other factors may also influence the appearance of heritage periods in X-men comics, and further research could explore these additional factors for a more comprehensive understanding.

```

## Discussion {#discussion}
Our report aimed to uncover the key variables that influence the total appearance count of X-Men characters. After conducting our analysis, we learned that the more profitable an X-Men character is, the more they're going to appear in X-Men comic-book issues. 
Did the data behave as you expected?
The data did not behave as we expected. 
Is the result statistically significant? How large is the effect? Would you consider it practically significant? How are you deciding? (Do you need to bring in any subject-matter knowledge here? For example 2 goals in hockey is a lot, but 2 points in basketball is a close game!)
Strengths (don’t restate but interpret) of data collection and analysis methods
Can you draw cause-and-effect conclusions?
To what population are you willing to generalize?
We can generalize our results to all X-Men characters that appeared in comic book issues in the decades between 60's and 90's.
How does your study contribute to the existing knowledge base and the larger context?
Critically assess any limitations or potential biases and how they impacted the interpretation of your findings.
Note any deficiencies or unexplained data.
Any issues with the analysis methods used?
Any other potential sources of error or biases?



This analysis aimed to identify the key factors that influence the total appearance count of X-Men characters in comic book issues from the 1960s to the 1990s. The results suggest that a character's profitability, as measured by their estimated monetary value, is positively correlated with their total number of appearances. This finding aligns with expectations, as publishers are likely to feature popular and lucrative characters more frequently to capitalize on their success and appeal to readers.
The effect of profitability on appearance count was found to be statistically significant, with a moderate to large effect size based on the regression analysis. This indicates that the relationship is unlikely due to chance and has practical significance in understanding character prominence. However, it is important to consider that while profitability is a strong predictor, it is not the sole determinant of appearance frequency. Other factors, such as narrative importance, character development, and creative decisions, also play a role in shaping character appearances.
The strengths of this study lie in the comprehensive dataset, which covers a substantial period of X-Men history and includes a diverse range of characters. The use of multiple statistical methods, including chi-square tests, linear regression, and ANOVA, allows for a thorough examination of the relationships between variables. However, the observational nature of the data limits the ability to draw definitive cause-and-effect conclusions. While profitability and appearance count are correlated, it is possible that other unmeasured variables are influencing this relationship.
The findings of this study can be generalized to the population of X-Men characters featured in comic books from the 1960s to the 1990s. However, caution should be exercised when extending these conclusions to other comic book franchises or time periods, as the dynamics and factors influencing character appearances may differ.
This research contributes to the existing knowledge base by providing empirical evidence for the link between character profitability and appearance frequency in the context of X-Men comics. It highlights the importance of considering financial factors when analyzing character prominence and opens avenues for further research on the interplay between commercial success and creative decisions in the comic book industry.
Limitations of this study include the potential for selection bias, as the dataset only includes a subset of X-Men characters and may not fully represent the entire franchise. Additionally, the reliance on estimated monetary values as a proxy for profitability may not capture all aspects of a character's financial success or appeal. Future research could explore alternative measures of profitability and expand the scope of the analysis to include more characters and a wider range of publication years.
In conclusion, this study reveals that profitability is a significant factor influencing the total appearance count of X-Men characters in comic books from the 1960s to the 1990s. While this finding provides valuable insights into character prominence, it is important to consider the limitations and potential biases of the analysis when interpreting the results. Further research is needed to fully understand the complex dynamics shaping character appearances in the comic book industry.
 
## Conclusion {#conclusion}

## References {.unnumbered}
