---
title: "Final Report"
author: 
  - name: Warren Woo
    affil-id: 1
  - name: Zachary Lakin
    affil-id: 1
  - name: Justin Mai
    affil-id: 1
  - name: Daniel Tsemekhman
    affil-id: 1
affiliations:
  - id: 1
    name: Department of Statistics, Cal Poly - SLO
date: last-modified
date-format: "5/31/24"
format: 
  html:
    self-contained: true
    code-tools: true
    toc: true
editor: source
execute: 
  error: true
  echo: false  
  message: false
  warning: false
abstract: |
Exploring the spotlight in superhero sagas: What determines which X-Men characters stand out in the pages of Marvel Comics? This study delves into the factors influencing the appearance frequency of X-Men characters in comic books published from 1963 to 1992. Motivated by the intriguing variance in character prominence, we analyze a dataset encompassing 26 X-Men, examining 44 variables per character, including total appearances, demographic details, and estimated financial values. Employing statistical techniques such as linear regression, we identify the significant predictors of visibility, with a particular focus on a character’s estimated monetary value. Our findings shed light on the complex dynamics of narrative and market influences in the X-Men series, offering insights into how profitability might shape character development in serialized media.
---

```{r}
#| label: setup
MutantData <- read.csv(
             "/Users/warre/Desktop/STAT365/MutantWithGenders.csv")
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(kableExtra)
library(forcats)


mean_ignore_zero <- function(x) {
  mean(x[x != 0], na.rm = TRUE)
}

average_values <- MutantData %>%
  group_by(Gender) %>%
  summarise(
    Avg_heritage_60s = mean_ignore_zero(TotalValue60s_heritage),
    Avg_ebay_60s = mean_ignore_zero(TotalValue60s_ebay),
    Avg_heritage_70s = mean_ignore_zero(TotalValue70s_heritage),
    Avg_ebay_70s = mean_ignore_zero(TotalValue70s_ebay),
    Avg_heritage_80s = mean_ignore_zero(TotalValue80s_heritage),
    Avg_ebay_80s = mean_ignore_zero(TotalValue80s_ebay),
    Avg_heritage_90s = mean_ignore_zero(TotalValue90s_heritage),
    Avg_ebay_90s = mean_ignore_zero(TotalValue90s_ebay)
  ) %>%
  gather(decade, value, starts_with("Avg")) %>%
  mutate(
    Source = ifelse(grepl("heritage", decade), "Heritage", "eBay"),
    Decade = as.numeric(sub(".*?(\\d+).*", "\\1", decade))
  )

MutantData <- MutantData %>%
  mutate(GrandTotalValue = TotalValue_heritage + TotalValue_ebay)
MutantData <- MutantData %>%
  mutate(
    decadeintroduced = case_when(
      X60s_Appearance_Percent != "0.00%" ~ "1960s",
      X70s_Appearance_Percent != "0.00%" ~ "1970s",
      X80s_Appearance_Percent != "0.00%" ~ "1980s",
      X90s_Appearance_Percent != "0.00%" ~ "1990s",
      TRUE ~ "Unknown"
    ),
    X90s_Appearance_Percent = as.numeric(sub("%", "", X90s_Appearance_Percent))
  )
```


## Introduction {#intro}
The X-Men, a cornerstone of Marvel Comics introduced in 1963, showcases a variety of mutants with unique abilities, yet their visibility in the comic series varies significantly. This study investigates what drives the prominence of certain characters over others, analyzing data from 1963 to 1992 to uncover the underlying factors influencing their appearance frequency.

Using a dataset compiled by Anderson Evans from the Marvel Fandom Database, we examine variables such as total appearances, appearance percentages by decade, estimated monetary values, and gender. The findings, based on rigorous statistical analysis, aim to offer a deeper understanding of the dynamics shaping narrative decisions in the X-Men comics. This report will begin by providing an overview of the data set and variables, followed by a detailed description of the statistical methods used. The results section will present the findings in the form of graphs and tables, and the discussion will interpret these results as well as discuss any issues or limitations with the data. Finally, the conclusion will summarize the key takeaways and suggest future directions for research on factors influencing X-Men character appearances.


## Data and Methods

### Data Collection

### Statistical Analysis

## Results {#results}

```{r}
#| label: T
#| fig-cap: "Table 1: The table displays all of the X-Men members total issues, ordered from greatest to least"
#| fig-align: center
#| fig-pos: "H"
library(kableExtra)
tabledata <- MutantData

tabledata <- tabledata[order(tabledata$TotalIssues, decreasing = TRUE), c("Member", "TotalIssues", "Gender"), drop = FALSE]


color.male = which(tabledata$Gender == 'M')
color.female = which(tabledata$Gender == 'F')

kbl(tabledata, booktabs = T) %>%
  kable_styling(position = "center") %>%
  kable_styling(font_size = 20) %>%
  row_spec(color.male, color = "blue") %>%
  row_spec(color.female, color = "pink1")
```
With a total of 26 X-Men members to go off of, 15 being male and 11 being female, the mean amount of total issues was 85.27. The mean for males was 94.73 and the mean for females was 72.36. The highest male was Scott Summers with 197 appearances and the lowest male was John Proudstar with only 3 appearances. The highest female was Ororo Munroe with 190 appearances and the lowest females was Remy Lebeau with 17 appearances.

After running a two-sample t-test, with a p-value of .365, we concluded that there was no significant different in total issue means between the two genders. Despite the fact that 6 out of the top 8 X-Men with the highest amount of total issues are males (Table 1), we concluded that there was not a significant difference in means. A 95% confidence interval of (-72.41, 27.67) would also further support this claim. Our degrees of freedom was 23.44. So in conclusion we that gender does not play a signficant role in determining the appearance rate of a X-Men.
 

```{r}
#| label: f
#| fig-cap: "Figure 1: This graph shows the variability in total issues for males and females X-Men characters. Each dot is one member."
#| fig-align: center
#| fig-pos: "H"


gender_boxplot <- ggplot(MutantData, aes(x = Gender, y = TotalIssues, color = Gender)) +
  geom_boxplot(aes(fill = Gender), alpha = 0.5) +
  geom_jitter(position = position_jitter(0.2), size = 1, alpha = 0.7) +
  labs(
    title = "Variability in Total Issues by Gender", 
    x = "Sex", 
    y = "Total Issues", 
    color = "Sex", 
    legend = "Sex",
    fill = "Sex"
  )  +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.spacing.y = unit(0, "mm"), 
    panel.border = element_rect(colour = "black", fill = NA),
    aspect.ratio = 1, 
    axis.text = element_text(colour = 1, size = 12),
    legend.background = element_blank(),
    legend.box.background = element_rect(colour = "black"),
    plot.caption = ggtext::element_markdown(
      size = 10,
      face = "italic",
      lineheight = 1.2,
      hjust = 0.5,
      vjust = 0
    )
  ) + 
  scale_color_manual(values = c("pink", "lightblue")) +
  scale_fill_manual(values = c("pink", "lightblue"))

print(gender_boxplot)

```
```{r}
#| label: h
#| fig-cap: "Figure 2: This graph displays the average appearance rate in the 90s,  categorized by the decade each character made their debut in."
#| fig-align: center
#| fig-pos: "H"

MutantData <- MutantData %>%
  mutate(
    decadeintroduced = case_when(
      X60s_Appearance_Percent != "0.00%" ~ "1960s",
      X70s_Appearance_Percent != "0.00%" ~ "1970s",
      X80s_Appearance_Percent != "0.00%" ~ "1980s",
      X90s_Appearance_Percent != "0.00%" ~ "1990s",
      TRUE ~ "Unknown"
    ),
    X90s_Appearance_Percent = as.numeric(sub("%", "", X90s_Appearance_Percent))
  )

Data_90s <- MutantData %>%
  filter(decadeintroduced != "Unknown") %>%
  select(decadeintroduced, Gender, X90s_Appearance_Percent)

appearance_data <- Data_90s %>%
  gather(key = "Decade", value = "AppearanceRate", X90s_Appearance_Percent)

average_rate <- appearance_data %>%
  group_by(decadeintroduced, Gender) %>%
  summarise(avg_appearance_rate = mean(AppearanceRate, na.rm = TRUE))

combined_rate <- appearance_data %>%
  group_by(decadeintroduced) %>%
  summarise(avg_appearance_rate = mean(AppearanceRate, na.rm = TRUE)) %>%
  mutate(Gender = "Combined")

combined_data <- bind_rows(average_rate, combined_rate)

decadeplot <- ggplot(combined_data, aes(x = decadeintroduced, y = avg_appearance_rate, fill = decadeintroduced)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Average Appearance Rate (In the 90s) by Decade Introduced",
    x = "Decade Introduced",
    y = "Average Appearance Rate (%)",
    fill = "Decade X-Men was Introduced"
  ) + 
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 80)) 

print(decadeplot)
```
```{r}
#| label: p
#| fig-cap: "Figure 3: Every dot is an X-Men, graphed by their total value, between both Ebay and Heritage, and total issues"
#| fig-align: center
#| fig-pos: "H"


scatter_plot <- ggplot(MutantData, aes(x = TotalIssues, y = GrandTotalValue)) +
  geom_point() +
  geom_smooth(method ="lm", se = FALSE, linetype = "solid", color = "navy") +
  labs(title = "Total Value vs Total Issues",
    x = "Total Issues",
    y = "Total Value ($)"
  ) + theme_bw() + theme(plot.title = element_text(hjust=0.5)) +
  scale_y_continuous(labels = scales::dollar)

print(scatter_plot)

```
Based on Figure 1, Figure 2, and Figure 3, the only variable to have correlation with appearance rate or total total issues was the characters total value. Based off of Table 1 and Figure 1, the gender of a character did not result in having a significant impact on total issues. In Figure 2 we see that there is no discernible pattern when looking at the average appearance rate between the decades. Despite the 1970s group having the highest average appearance rate, the 1960s and 1980s had the 3rd and 4th highest average appearance rate which does not show a pattern. The decade in which a character was introduced in does not seem to have an impact on their appearance rate. In Figure 3 however, we can seen that the graph shows a general positive correlation between appearances and total value, with outliers of high-value characters with few appearances and a dense clustering of many characters with lower appearances and total values. Meaning that characters with a higher amount of total appearances tend to have a higher total value, and characters with a lower amount of total appearances tend to have a lower total value. After running a regression analysis, we find that the p-values for both the intercept and the slope are less than .05, indicating that the total value of character is a strong indicator of total issues. 

## Discussion {#discussion}
Our report aimed to uncover the key variables that influence the total appearance count of X-Men characters. After conducting our analysis, we learned that the more profitable an X-Men character is, the more they're going to appear in X-Men comic-book issues. This analysis aimed to identify the key factors that influence the total appearance count of X-Men characters in comic book issues from the 1960s to the 1990s. The results suggest that a character's profitability, as measured by their estimated monetary value, is positively correlated with their total number of appearances. This finding aligns with expectations, as publishers are likely to feature popular and lucrative characters more frequently to capitalize on their success and appeal to readers. The effect of profitability on appearance count was found to be statistically significant, with a moderate to large effect size based on the regression analysis. This indicates that the relationship is unlikely due to chance and has practical significance in understanding character prominence. However, it is important to consider that while profitability is a strong predictor, it is not the sole determinant of appearance frequency. Other factors, such as narrative importance, character development, and creative decisions, also play a role in shaping character appearances. The strengths of this study lie in the comprehensive dataset, which covers a substantial period of X-Men history and includes a diverse range of characters. However, the observational nature of the data limits the ability to draw definitive cause-and-effect conclusions. While profitability and appearance count are correlated, it is possible that other unmeasured variables are influencing this relationship. The findings of this study can be generalized to the population of X-Men characters featured in comic books from the 1960s to the 1990s. However, caution should be exercised when extending these conclusions to other comic book franchises or time periods, as the dynamics and factors influencing character appearances may differ. This research contributes to the existing knowledge base by providing empirical evidence for the link between character profitability and appearance frequency in the context of X-Men comics. It highlights the importance of considering financial factors when analyzing character prominence and opens avenues for further research on the interplay between commercial success and creative decisions in the comic book industry.Limitations of this study include the potential for selection bias, as the dataset only includes a subset of X-Men characters and may not fully represent the entire franchise. Additionally, the reliance on estimated monetary values as a proxy for profitability may not capture all aspects of a character's financial success or appeal. Future research could explore alternative measures of profitability and expand the scope of the analysis to include more characters and a wider range of publication years.In conclusion, this study reveals that profitability is a significant factor influencing the total appearance count of X-Men characters in comic books from the 1960s to the 1990s. While this finding provides valuable insights into character prominence, it is important to consider the limitations and potential biases of the analysis when interpreting the results. Further research is needed to fully understand the complex dynamics shaping character appearances in the comic book industry.
 
 
## Conclusion {#conclusion}
Ultimately, our data suggests that if you're a profitable X-Men character, you get put in more X-Men comic book issues. Within the confines of the dataset, outside of monetary value, there aren't any other statistically significant contributing factors to a X-Men’s total appearances. We can only generalize our results to other X-Men comic book characters, because the data only looked at X-Men characters. Drawing a conclusion to other Marvel characters would be unwise because our sample isn’t representative of the Marvel population as a whole. In other words, our research has low generalizability; our results are only representative of our sample containing X-Men characters. Furthermore, we cannot infer causality because there simply exists many more variables that contribute to an X-Men’s total appearances, which are not included in the dataset; variables such as: an X-Men character’s writer, an X-Men’s amount of media appearances outside of the comics, and how many plots/storylines does an X-Men character appear in. For those interested and curious - these variables necessitate more investigation. Theoretically, there are infinitely many variables that could contribute to an X-Men’s number of appearances in comic book issues, and if one was extremely curious, they could run a study including some of these variables to continue our research and attempt to answer the question: “What drives the prominence of certain X-Men characters over others? What are the underlying factors influencing their appearance frequency?” By furthering our research, we believe that these determined researchers will discover some groundbreaking results that could change the way comic book writers write their characters.

## References {#references}

Marvel database. Marvel Fandom Database. (n.d.). https://marvel.fandom.com/wiki/Marvel_Database 

Evans, A. (2023, June 16). Mutant Moneyball: A Data Driven Ultimate X-Men. Rally. https://rallyrd.com/mutant-moneyball-a-data-driven-ultimate-x-men/ 

The R project for statistical computing. R. (n.d.). https://www.r-project.org/ 

Zhu H (2024). _kableExtra: Construct Complex Table
  with 'kable' and Pipe Syntax_. R package version
  1.4.0,
  <https://CRAN.R-project.org/package=kableExtra>.


Wickham H, Averick M, Bryan J, Chang W, McGowan LD,
  François R, Grolemund G, Hayes A, Henry L, Hester
  J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller
  K, Ooms J, Robinson D, Seidel DP, Spinu V,
  Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H
  (2019). “Welcome to the tidyverse.” _Journal of
  Open Source Software_, *4*(43), 1686.
  doi:10.21105/joss.01686
  <https://doi.org/10.21105/joss.01686>.

H. Wickham. ggplot2: Elegant Graphics for Data
  Analysis. Springer-Verlag New York, 2016.

Wilke C, Wiernik B (2022). _ggtext: Improved Text
  Rendering Support for 'ggplot2'_. R package version
  0.1.2, <https://CRAN.R-project.org/package=ggtext>.

Wickham H (2023). _forcats: Tools for Working with
  Categorical Variables (Factors)_. R package version
  1.0.0,
  <https://CRAN.R-project.org/package=forcats>.

Wickham H, Vaughan D, Girlich M (2024). _tidyr:
  Tidy Messy Data_. R package version 1.3.1,
  <https://CRAN.R-project.org/package=tidyr>.

Wickham H, François R, Henry L, Müller K, Vaughan D
  (2023). _dplyr: A Grammar of Data Manipulation_. R
  package version 1.1.4,
  <https://CRAN.R-project.org/package=dplyr>.

